### 2020-01-17 manuscript/supplement.tex
added supplement

\documentclass[aps,rmp,onecolumn]{revtex4}
%\documentclass[a4paper,10pt]{scrartcl}
%\documentclass[aps,rmp,twocolumn]{revtex4}

\usepackage[utf8]{inputenc}
\usepackage{amsmath,graphicx}
\usepackage{color}
%\usepackage{cite}

\newcommand{\bq}{\begin{equation}}
\newcommand{\eq}{\end{equation}}
\newcommand{\bn}{\begin{eqnarray}}
\newcommand{\en}{\end{eqnarray}}
\newcommand{\Vadim}[1]{{\color{blue}Vadim: #1}}
\newcommand{\Richard}[1]{{\color{red}Richard: #1}}
\newcommand{\gene}[1]{{\it #1}}
\newcommand{\mat}[1]{{\bf #1}}
\newcommand{\vecb}[1]{{\bf #1}}
\newcommand{\abet}{\mathcal{A}}
\newcommand{\eqp}{p}
\newcommand{\pc}{c}
\newcommand{\LH}{\mathcal{L}}
\newcommand{\lh}{\ell}
\pdfinfo{%
  /Title    ()
  /Author   ()
  /Creator  ()
  /Producer ()
  /Subject  ()
  /Keywords ()
}

\begin{document}

\title{Supplement: Potential and limits of site-specific substitution model inference.}
\author{Vadim Puller$^{1,2}$ Pavel Sagulenko$^{3}$, Richard A.~Neher$^{1,2}$}
\affiliation{
$^{1}$Biozentrum, University of Basel, Klingelbergstrasse 50/70, 4056 Basel, Switzerland\\
$^{2}$SIB Swiss Institute of Bioinformatics, Basel, Switzerland\\
$^{3}$Max Planck Institute for Developmental Biology, 72076 T\"ubingen, Germany}

\date{\today}
\maketitle

For a given set of sequences $\{\vec{s_k}\}$ consisting of the alignment at the tips and the ancestral sequences, the log-likelihood is given by
\begin{equation}
	\label{eq:logLH}
	\lh(\{\vec{s}\} |T, \mat{Q}) = \sum_a \left[\log(\eqp^a_{s^{a}_0}) + \sum_{\beta\in T} \log\left( e^{Q^at_\beta} \right)_{s^a_c, s^a_p}\right]
\end{equation}
where $\mat{Q}$ is the site specific substitution model consisting of rates, frequencies, and a symmetric transition matrix
\begin{eqnarray}
Q^{a}_{ij} &=& \mu^{a}\eqp^{a}_{i} W_{ij} \textrm{ for } i\neq j,\nonumber \\
Q^{a}_{ii} &=& -\sum_k Q^{a}_{ki} \ .
\label{eq:Qij_supp}
\end{eqnarray}

At the maximum, the derivatives of $\lh$ with respect to each of the model parameters should vanish when averaged over all internal site assignments.
The derivative with respect to $\mu^b$ was is presented in Eq.~12 of the main text.
Here, we present the corresponding derivations for $\eqp^b_n$ and $W_{nm}$.

To facilitate the algebra below, we summarize the short branch length assumption and the associated linear approximations here:
\begin{equation}
	\left( e^{Q t}\right)_{ij} = \delta_{i,j} + \sum_{m=1}^\infty \frac{t^m}{m!}\left( Q^m\right)_{ij} \approx
	\begin{cases}
	1 - t\mu^b \sum_{k\neq i} \eqp_k^b W_{ki} & i=j \\
	t\mu^b \eqp^b_i W_{ik} & i\neq j
	\end{cases}
\end{equation}
The derivatives of $\left( e^{Q t}\right)_{ij}$ with respect to the different model parameters are therefore
\begin{eqnarray}
	\frac{d}{d\mu^b} \left( e^{Q^a t}\right)_{ij} &\approx & \delta_{ab}
	\begin{cases}
	-t\sum_{k\neq i} \eqp_k^b W_{ki} & i=j \\
	t \eqp^b_i W_{ik} & i\neq j
	\end{cases} \\
	\frac{d}{d\eqp_n^b} \left( e^{Q^a t}\right)_{ij} &\approx & \delta_{ab}
	\begin{cases}
	-t\mu^b (1-\delta_{i,n})W_{ni} & i=j \\
	 t\mu^b \delta_{i,n}W_{nk} & i\neq j
	\end{cases}\\
	\frac{d}{dW_{mn}} \left( e^{Q^a t}\right)_{ij} &\approx &
	\begin{cases}
	-\delta_{n,j}(1-\delta_{m,i})t \mu^a \eqp^a_m  & i=j \\
	 \delta_{m,i}\delta_{n,j} t\mu^a \eqp^a_m & i\neq j
	\end{cases}
\end{eqnarray}
Since the matrix $W_{ij}$ is symmetric, the derivative with respect to $W_{ij}$ and $W_{ji}$ are the same.
We will account for this symmetry below explicitly and refrain from carrying the symmetric term through lengthy derivations.
Note that the derivatives with respect to $W_{ii}$ vanish as required.

\section*{Update rule for rates $\mu^b$}
Only site $b$ will contribute to the derivative of $\lh$ with respect to $\mu^b$.
We will therefore drop all indices referring to the position and re-instantiate the index later.
As discussed in the main text, we will separate the sum over branches $\beta$ into those where sequence is the same of parent and child ($s_c=s_p$) and those where the sequence differs ($s_c\neq s_p$).
\begin{equation}
\begin{split}
	\frac{d}{d \mu} \lh (\{\vec{s}\} |T, \mat{Q}) &
	 \approx  -\sum_{\beta\in T,s_c=s_p} \frac{t_\beta \sum_{k\neq s_c} \eqp_{k} W_{k s_c}}{1 - t_\beta \mu \sum_{k\neq s_c} \eqp_k W_{ks_c}}
	 + \sum_{\beta\in T,s_c\neq s_p} \frac{t_\beta \eqp_{s_c} W_{s_cs_p}}{t_\beta \mu \eqp_{s_c} W_{s_c s_p}}\\
	 & \approx  -\sum_{\beta\in T,s_c=s_p} t_\beta \sum_{k\neq s_c} \eqp_{k} W_{k s_c} + \sum_{\beta\in T,s_c\neq s_p} \frac{1}{\mu} = - \sum_{j, k\neq j} \eqp_k W_{kj} \tau_j + \sum_{i\neq j} n_{ij}/\mu
\end{split}
\end{equation}
Here, $\tau_j$ is the sum of all branch lengths on which $s_c=j$ and $n_{ij}$ is the number of times the sequence changes from $j$ to $i$ somewhere along the tree.
Setting this expression to 0 and re-instantiating the site index $b$, we get
\begin{equation}
	\mu^b = \frac{\sum_{i\neq j} n^b_{ij}}{\sum_{j, k\neq j} \eqp^b_k W_{kj} \tau^b_j}
\end{equation}

\section*{Update rule for frequencies $\eqp_i^b$}
As above, only site $b$ will contribute to the derivative of $\lh$ with respect to $\eqp^b_n$.
We will therefore drop all indices referring to the position and re-instantiate the index later.
\begin{equation}
\begin{split}
	\frac{d}{d \eqp_n} \lh (\{\vec{s}\} |T, \mat{Q}) & \approx \frac{\delta_{s_0,n}}{\eqp_n}
	-\sum_{\beta\in T,s_c=s_p} \frac{t_\beta\mu (1-\delta_{n,s_c})W_{n s_c}}{1 - t_\beta\mu \sum_{k\neq s_c} \eqp_k W_{ks_c}}
	 + \sum_{\beta\in T,s_c\neq s_p} \delta_{n,s_c}\frac{t_\beta\mu W_{ns_p}}{t_\beta \mu \eqp_{s_c} W_{s_c s_p}}\\
	 & \approx 	\frac{\delta_{s_0,n}}{\eqp_n}-\sum_{\beta\in T,s_c=s_p} t_\beta\mu (1-\delta_{s_c, n})W_{s_c n}
	 + \sum_{\beta\in T,s_c=n, s_c\neq s_p} \frac{1}{\eqp_{n}}\\
	 & = 	\frac{\delta_{s_0,n}}{\eqp_n}-\sum_{j\neq n} \tau_j \mu W_{j n}
	 + \sum_{j\neq n}\frac{n_{nj}}{\eqp_{n}}
\end{split}
\end{equation}
Reinstantiating the position index $b$, we find
\begin{equation}
	\eqp^b_n = \frac{\delta_{n,s^b_0}+\sum_j n^b_{nj}}{\sum_{j\neq n} \mu W_{nj}\tau^b_j}
\end{equation}

\section*{Update rule for the transition matrix $W_{ij}$}
Since $W_{ij}$ does not depend on the position, the derivative of $\lh$ with respect to $W_{ij}$ depends on all sites in the alignment.
\begin{equation}
\begin{split}
	\frac{d}{d W_{mn}} \lh (\{\vec{s}\} |T, \mat{Q}) & \approx
	-\sum_a\sum_{\beta\in T,s^a_c=s^a_p} \frac{\delta_{n,s^a_p}(1-\delta_{m,s^a_c})t_\beta \mu^a \eqp^a_m}{1 - t_\beta\mu \sum_{k\neq s^a_c} \eqp_k W_{ks^a_c}}
	 + \sum_{\beta\in T,s_c\neq s_p} \delta_{m,s_c}\delta_{n,s_p} \frac{t_\beta \mu^a \eqp^a_m}{t_\beta \mu^a \eqp^a_{m} W_{m n}}\\
	 &\approx -\sum_a\sum_{\beta\in T,s^a_c=s^a_p}  \delta_{n,s^a_p}(1-\delta_{m,s^a_c})t_\beta \mu^a \eqp^a_m + \frac{1}{W_{mn}}\sum_a n^a_{mn}\\
	 & = \sum_a \tau^{a}_{n} \mu^a \eqp^a_m  + \frac{1}{W_{mn}}\sum_a n^a_{mn}
\end{split}
\end{equation}
Identifying $W_{nm}$ and $W_{mn}$ then results in the following symmetric expression
\begin{equation}
	\frac{d}{d W_{mn}} \lh (\{\vec{s}\} |T, \mat{Q}) \approx \sum_a \mu^a (\tau^{a}_{n} \eqp^a_m + \tau^{a}_{m} \eqp^a_n)  + \frac{1}{W_{mn}}\sum_a (n^a_{mn}+ n^a_{nm})
\end{equation}
Setting this expression to 0 and solving for $W_{ij}$ we find
\begin{equation}
	W_{mn} = \frac{\sum_a n^a_{mn}+ n^a_{nm}}{\sum_a \mu^a(\tau^{a}_{n} \eqp^a_m + \tau^{a}_{m} \eqp^a_n)}
\end{equation}

\section*{Implementation of the constraint $\sum_i \eqp_i^b=1$}
Equilibrium frequencies need to be normalized to one and constraints of this nature are typically accounted for by Lagrange parameters
\begin{equation}
	\frac{d}{d \eqp_n} \left[\lh (\{\vec{s}\} |T, \mat{Q}) + \lambda^b \sum_i \eqp^b_i \right] = 0
\end{equation}
which results in a modified extremal condition
\begin{equation}
	0 = \delta_{s^b_0,n}-\eqp_{n}\sum_{j\neq n} \tau^b_j \mu^b W_{j n} + \sum_{j\neq n} n^b_{nj} + \eqp^b_{n}\lambda^b
\end{equation}
Summing this condition over $n$, we find
\begin{equation}
	0 = 1 - \sum_{i,j; i \neq n}  \mu^b \tau^b_j \eqp^b_{i}W_{j i} + \sum_{i,j; i\neq j} n^b_{ij} + \lambda^b
\end{equation}
At the extremum of $\mu^v$, we have $\sum_{i,j; i \neq n} \mu^b \tau^b_j \eqp^b_{i}W_{j i} = \sum_{i,j; i\neq j} n_{ij}$ and therefore $\lambda^b=-1$.
With the constraint $\sum_i \eqp_i^b=1$, the update rule therefore becomes
\begin{equation}
	\eqp^b_n = \frac{\delta_{n,s^b_0}+\sum_j n^b_{nj}}{1 + \sum_{j\neq n} \mu W_{nj}\tau^b_j}
\end{equation}
In practice it is often helpful to explicitly enforce normalization after each iteration.

\section*{Regularization}
In absence of data, i.e., when there are very few mutations at a particular site we need to regularize the problem.
To this end, we assume a Dirichlet prior of the form
\begin{equation}
\prod_i \left(\eqp^b_i\right)^c
\end{equation}
The derivative of the logarithm of this is $c/\eqp^b_i$ and adding this term to the update rule results in
\begin{equation}
	\eqp^b_n = \frac{\delta_{n,s^b_0} + c +\sum_j n^b_{nj}}{1 + qc + \sum_{j\neq n} \mu W_{nj}\tau^b_j}
\end{equation}
where $q$ is the size of the alphabet.
In an analogous fashion, we add a weak regularization to the update rules of $\mu^b$ and $W_{ij}$.

\section*{Summation over unobserved states}
So far, we have derived conditions that maximize the likelihood as a function of parameters of the substitution model for a fixed assignment of sequences at internal nodes.
In practice, the internal nodes are unknown and can either be fixed to the most-likely state or summed over.
Since the approximate maximum likelihood conditions are linear, this average over unknown ancestral states simply amounts to replacing $\tau^b_i$ and $n^b_{ij}$ by their averages which can be efficiently computed by standard recursive algorithms.

\end{document}